Sketch Recogniton using Particle Swarm Algorithm

### 2009-04-12 trunk/draft/chapterSurvey.tex
12-4

\chapter{Literature Survey}

\section{Introduction }
\label{sec:Introduction}

The researchers had divided the sketch recognition problem into image based and AI Based problem. The next section will discuss the methods used by current sketch understanding system. Firstly, in section \ref{sec:FirstStepSymbolRecognition} Symbol recognition research will be investigated and divided into image based and AI based methods.

 In section \ref{sec:SecondStep:SketchUnderstanding}, the understanding techniques will be demonstrated. Understanding mainly will handle knowledge representation methods, handling uncertainty and finally how researcher in sketch reduced the sketch search space.

\section{First Step: Symbol Recognition}
\label{sec:FirstStepSymbolRecognition}

To understand a drawing we need to recognize every symbol in it. Hand-drawn sketches need more processing before it is passed to recognitions algorithms.

 Firstly, we have to determine how symbols are defined or grouped. Some System implies some constrain on user to defined symbol. In\cite{gestureexample12,aideddesgin22,sketchinginterfaces2,sketchinginterfaces2,cognitivesketch18,ewhitboard9}for example the user is constrained to draw each symbol using only one stroke which is not only unnatural but also constrains the symbols that can be drawn. More recent systems force the user to state when symbol is finished by pausing a period of time or clicking a certain button in the interface. Only few systems permit the user to freely draw strokes and symbols in any order \cite{Vibratory8,visualpattern43,Cali63,statisticalparsing26,physicalmeaning6}. Those systems must employ a method to detect when a user starts drawing a symbol and when he end.

Currently symbol recognition is both image based and feature based or AI-based. In image based recognition symbol is recognized based on there features, structure, some image computed feature . The AI methods in recognizing symbols mainly imply more time and processing as for the search and rules applied but claims to be more accurate .

\subsection{Pre processing }
\label {sec:preprocessing}
 A stroke is all the points that lie along the path of the user pen from the moment of his pen down to pen up events. Not all those points are important even more, there may be redundancies unwanted points. Therefore, those strokes will need some preprocessing.

The stroke that the user draws are mainly sequences of line segments or/and curves which the any system need to divide them to recognize them easily. The main problem is to do so without constraining the user drawing style. The majority of researchers first use a preprocessing process to identify those segments and the critical points from the user strokes.  
Beside points location information (x, y positions) researcher made uses of the extra information from using digital ink. Some as \cite {mulitstroke5,polygonfeedback31}  used user speed information to detect critical points where it is stated that user slow down at those critical points. Other used curvature and directions data. There was also mentioning using the pressure information.  Thought some \cite {MilitaryCOA37,computationalmodel16,polygonfeedback31}  debated on the importance of the extra information and weather the temporal and speed information is significant.  \cite{polygonfeedback31}  used the timestamp for each point and compute time difference between every two successive point to help him finding corners.
\cite{Overview36} uses the speed and direction information but they use on it a hybrid average based filtering to filter noise and unwanted false data from appearing in the data. Finally, they use that information to get a poly line approximating.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/dirspeedgraph.eps}
	\caption[Direction and Speed]{direction and speed graphs of the rectangle stroke the minima in the graph show the location of the corners of the rectangle.}
	\label{fig:dirspeedgraph}
\end{figure}

\cite{Mathpad46}Handle rotations invariant using polar transformation then shifting to make it more reliable when template matching. However, they state that the only problem to there system is that the points near the centroid affect the polar transformations which will effect the matching. (To solve it they say that loosing the weight of the pixel near the centroid will improve classification).

The next sections will demonstrate firstly the noise removal algorithms and techniques. Then next sections will show the most used critical points detection algorithms along with the segment division algorithms.

  In\cite{sketchRead29} they divided user strokes into geometric primitives based on some critical points they computes.  After that, they either cluster segment to form a symbol   or directly apply some kind of basic symbol recognition leaving the compound object for next processing step .

From the image-based view the symbol is a sequences of line segments or/and curves. To do so without constraining the user drawing style we have to easily divide the user drawn strokes.

%\begin{figure}
%	\includegraphics[scale=0.6]{../../neededfiles/Figures/shapesegmentation.eps}
%	\caption[segmented shapes]{example of Segmented shapes}
%	\label{fig:shapesegmentation}
%\end{figure}

\subsubsection{Removing noise}
\label{sec:RemovingNoise}

Due to the sloppiness of the user, his stroke may need a lot of noise removal before any algorithm can extract its features. For example, when a user draw sketches he may over trace the stroke more than one time. Human perception can easily differentiate what exactly is drawn. Nevertheless, for an AI program this is more difficult problem. The user may also overshoot see \textit{Fig\ref{fig:Overshootandundershoot}} or undershoot while drawing intersecting segments. 
\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/Overshootandundershoot.eps}
	\caption[Neat and Sloppy symbols] {The Figure show a sloppy strokes with under shoot and overshoot problems and the neat version of the symbol.}
	\label{fig:Overshootandundershoot}
\end{figure}

 \cite {overtraced24} tried to solve the over traced strokes. they used a thinning algorithm to move points perpendicularly towards the regression line from the formed from fitting line of neighbour's into line or circle or ellipse which is calculated with least square fit for the data. The next step is selecting near neighbours and connection then which gives a serious of linked edges which a greedy link edge algorithm this produce a single stroke instead of over traced see \textit{Fig [\ref {fig:overstroked}]}.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/overstroked.eps}
	\caption{Example of an overstroked stroke}
	\label{fig:overstroked}
\end{figure} 
Some systems as in \cite{SmartSketch56} where the noise reduction is done in two steps the first step is the polygonal approximation to remove non critical points in the user stroke. The process is mainly removing the redundant points that are along the path of stroke. Due to the shaky nature of user stroke, there are additional end-points at the stroke segments. Secondly, the next step will be applying an Agglomerate point filtering.  Agglomerate filtering is according to \cite {statesurvey35} removing the undesired extra points that is produced due to the unsteady operations produced while lifting and touching tablet.  The results of those processes are passed to final end-point refinement to merge or remove any redundant points.

In \cite{meanshift10,domainindependent17}, Bo Yu uses mean shift procedure as a noise remover and preprocessing step to the recognition step. The mean shift procedure is simply shifting the point to the average of its neighborhoods so smoothing out the noise without changing the global structure of data.  This process is clamed to help find critical points easily after that.

In \cite {domainindependent17}    approximation step uses the speed, curvature, and direction of strokes drawn by user to determine the location of vertices to gain more resistance to noise that data may contain.

\subsubsection{Critical points and corners Detection}
\label{sec:CriticalPointsAndCornersDetection}

Locating the critical points is provided to the next process of the systems which in most cases segment detection algorithm as in \cite{Simusketch51} . Other systems after detecting the points the system directly uses those information to template matching.

Generally, the critical points along with the segment points of the path from one critical point to another are used to detect labeling of that segment. 
Line detection or multi line segments algorithm is found in \cite{polygonfeedback31} . Mostly after detecting the vertices if the segment cannot be labeled as arc it is defined as a line. Other mainly defines a threshold value of the slope if it increase than that it is counted as an arc.

Systems like \cite{physicalmeaning6} detect critical point for each stroke then the final point is merged with the first if they are in range defined by threshold. However, stroke may also intersect with other strokes drawn before therefore \cite{physicalmeaning6}  introduced a structure to detect such approximates he first define every stroke drawn as series point and labeled as line, arc or area. Then after labeling most strokes the system check for proximity on bases of line -line, line-end point, or line- area.

In   \cite {meanshift10,domainindependent17}   approximation step uses the speed, curvature, and direction of strokes drawn by user to determine the location of vertices to gain more resistance to noise that data may contain. They firstly analysis the curvature and speed data and use the mean of data as the threshold then use that threshold to separate data into regions and finally locating the maxima and minima in each region . The previous step generates a set of vertices and its certainties they use to choose best fit.

\subsubsection {Primitive stroke segmentation and clustering}
\label{sec:Primitivestrokesegmentationandclustering}

The most segmentation method is least square method in \cite {overtraced24}. Other systems  approximate a line segmented using the information from the direction curve where the curve is fitted with horizontal line. firstly the system compute least square fit for direction curve or the stroke point computed and check if deviation points is small if they pass then hypothesize al line segment  from the first point to the last one.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/rectanglerecog.eps}
	\caption{rectangle recognition}
	\label{fig:rectanglerecog}
\end{figure}

%//Other like \cite {tobeadd}  uses the relaxation process to try to relax parameter on a classification .//
 In \cite {meanshift10,domainindependent17}  Bo first uses the direction graph to check if the stroke can be approximated as a line or no. If the stroke can't be approximated as a line it is checked to be a circle or ellipse. the procedure is simply assume a circle classification then check this assumption with the user drawn stroke given a min error range.  
\cite {Fluid25} Trying to combine fast morphing and basic recognition for the sketch drawn as the user draw it. It tries to move the points of strokes drawn by the user draw into one of the known shapes (ex: box, lines and circle). It uses a relaxation and least square method for circles and uses the string of forces for finding best-fit box.

The noise removal and critical point detection is only a preprocessing step for the symbol classification algorithms. Some system implements a segment segmentation and clustering to help in the classification problem.

After the strokes are segmented \cite {sketchunderstanding1,HierarchicalParsing7,Simusketch51}  proposed a merging techniques that will merge strokes together into symbols to help classification. They use a method combine between divide and merge. They firstly divide the drawn sketches into section based on marker technique they used. After that they merge strokes if there bounding boxes intersects and then  merge generated cluster till no more merge can occur on one segment.

%Intension exteraciton in paper Intention54
% Check to add

%/*Mathpad46
%methion  check stroke clustering.  Simusketch51  and in 8 
%
%Bo \cite {tobeadd}   The ellipse or

\subsubsection{Image and Statistical Features calculation}
\label{sec:ImageandStatisticalFeaturescalculation}

Most system calculates the general image properties for the strokes. Most used features are bounding box, width, size, points along path. 
\cite {meanshift10,domainindependent17}  introduced a new feature which they named "Feature Area which" they used in primitive classification introduced before. They first make and assumption for classification and compute the difference. For example if they make assumption as a circle they first calculate the center of the circle then its radius based on the bounding box of the stroke. The area difference between the computed circle and the current stroke is the feature area. see \textit{Fig\ref{fig:featurearearc}}.

\begin{figure}

		\subfigure[Feature area of line] {\includegraphics[scale=0.7]{../../neededfiles/Figures/featureareline.eps}}

	\subfigure[Feature area of an arc]{	\includegraphics[scale=0.7]{../../neededfiles/Figures/featurearearc.eps}	}
			\caption[Feature area]{the shaded are shows the difference between the computed feature area and the assumption area. }
	\label{fig:featurearearc}
\end{figure}

 System that uses statistical classifiers use other features for each stroke. Features like center of gravity\cite{gestureexample12,sketchunderstanding1}, maximum length line, Zenerik moments \cite{zernike61}, number of lines, curvature angles, minimum enclosing rectangle\cite {incrmentintention41} , convex hull were calculated in \cite {Cali63}.  Also \cite {geometrydomain49} introduced a new feature which he named ink density that he used in the classification of the glyphs see \textit{Fig \ref{fig:inkdensity}}. 
\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/inkdensity.eps}
	\caption[Statistical Feature Example]{The figure shows the Ink Density feature that \cite{geometrydomain49} introduced. }
	\label{fig:inkdensity}
\end{figure}
All spatial relationships needed in the graph or template matching is computed after statistical features are calculated. Relations like intersections, containment, perpendicular, right-of, left-of and below are the most used relations concerning constructing graphs for symbols\cite {geometrydomain49,sketchinginterfaces2} .%stretchdiagrammer4,     i dont know where is this refrence now

\subsection{Image Based Classification}
\label{sec:Image Based Classification}

In the following sections, systems generally try to recognize the strokes user draws and give it a label based on there image features. The information collected and features computed in the previous step.
 \subsubsection{Low Level Analysis}
\label{sec:Low Level Analysis}

%
%Detecting object using low level analysis only 
%\cite {mulitstroke5} first detecting vertices and control points then fit these segment into line and arc using best fit and least square error classification.

%Check the DB

% description[4] after that it passes through the curve approximation into the Bezier curves
%
%\cite {statesurvey35} we need to fit and filter the classified shapes that is done using on of the following least square is the most frequently method used some use fuzzy logic to make circular fitting some use b-splines or ellipse and circle fitting. some after that do some  shape beatification or regularization.

The most used method in low level recognition is best fit and least square methods where they was used in most systems. Other systmes like \cite{polygonfeedback31} prefered to implement its own polygon recognition algorithm.

Bo et al. In \cite{meanshift10} introduced an algorithm to recognize line and arcs.  The system used the direction graph and fitted stroke points with a non horizontal line. So the recognizing is done using the following algorithm which is preformed before the mean shift procedure so not the destroy the direction graph .Firstly, if direction graph can be approximated using a line then it is recognized as a line. Otherwise the algorithm Bisect the line from the first and last  point of stroke and donate it as "M". Then calculate the intersection point of the line "M" and original stroke and donate it as "A" . Now using the three points first , last points of stroke and point A find the center of circle which passes through all of them which will donate the candidate arc. Finally , calculate the radius of the circle and the mean of distance  between center and each point of stroke and  compute the feature area of the stroke and the circle arc see \textit{Fig \ref{fig:featurearearc}}.
%//

\subsubsection{Template matching}
\label{sec:Template matching}

Template matching is the most used method to recognize strokes and symbols into classes.  Image based template stored on the database is matched with the one user draw to find the correct classification of the strokes drawn.  In most cases, the stroke template is stored as image bitmap \textit{Fig \ref{fig:template}}in the database when the user draws the symbol it is matched using classifiers. \cite {imagetrainable48,HierarchicalParsing7}  used four kind of classifiers to detect the matching template. He used two classifiers to measure similarity and two classifiers to measure dissimilarity then combine them to find best match of drawn stroke from the database.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/template.eps}
	\caption[Template matching]{Examples of image templates used to match symbols in \cite{imagetrainable48}. }
	\label{fig:template}
\end{figure}

To refine the matching procedure and create more robust classification \cite {imagetrainable48}  first transform the strokes into polar coordinate then match the transformed stroke with the templates in the database. They claim that enhance the classification procedure as the polar coordinate system due to its definition make the stroke more robust to transformation and rotation.

\cite{templatefrag21}  Introduce new way to match stroke temples. He store template as a sequence of lines and ellipse.  When the user draws a stroke the system, segment it into a sequence of lines and ellipse. The system then matches the stored sequences with computed one to find correct classification.

The main problem of template matching is that in most system the templates are hard coded inside the system and cannot be modified at runtime.

\subsubsection{Graph matching}
\label{sec:Graph matching}

Representing the symbol that the user draws as a graph transforms the classification problem into a graph-matching problem. In general, each user stroke is represented as a node of the graph and the relations between those nodes are represented as arc. To classify the symbols the graph constructed must be matched with the graphs in the database.

\cite{physicalmeaning6} uses spatial relationships for there graph matching. They first calculate spatial proximity relationships. Relations like intersections, containment, perpendicular, right-of, left-of and below are the most used relations concerning constructing graphs for symbols.

Other systems  first analysis the spatial relationship between glyphs (sequence of strokes) using a library of spatial relations and information calculated in the preprocessing step.  After that, they filter out important relation-using hierarchy of rules and type of geometrical constrain. All redundant constrains and relations are removed the final graphs is matched with the database. If more than one candidate is available the algorithm choose the best fit but if no matches found constrains are relaxed and strokes are passed again to classifier.

 Proposed a structure representation of the strokes in \cite{sketchunderstanding1,HierarchicalParsing7} . She added ratio of proximity of the relations between strokes. In addition, she used an error-driven, stochastic search algorithm for graph matching.

\subsubsection{Statistical methods}
\label{sec:Statisticalmethods}

The application of Statistical classifiers in recognizing shapes are widely used in sketch recognition. Statistical classifier requires training time before the recognition actually happen. Then first step of classification is features computation that is done in the previous step. Finally, the recognition of symbols will be based on the classifier and the training set used.

One of the first systems in sketch recognition is \citep{gestureexample12,aideddesgin22} algorithm to recognize ink gestures. He used Gaussian statistical classifier to classify any stroke user draw.  In \cite {gestureexample12}  he firstly compute statistical information about the user stroke then pass it to a Gaussian classifier that had been trained using 10-20 samples. The output of the classifier is classification of the stroke from one of the known classes.

Bayesian classifiers was used in \cite{Vibratory8} as a stroke classifier to recognize different shapes. The recognition is done using feature based statistical Gaussian distribution and Bayesian classifier techniques. The trainer use nine features extracted from segmented symbols (ex number of line segments, number of L intersections, ) all feature has a tolerance within. After all nine feature are computed a Gaussian statically model is constructed.

Some systems employ more than one statistical classifier to get better results for the recognition. For example \cite{zernike61} uses Neural network, SVM (support vector machine )and MMD. \cite{polygonfeedback31}
%the closed shaper recognizer which is done using classifiers (SVM , Rule based and ANN ) to make a 
%classification of triangle, rectangle or ellipse

\subsubsection{Other hybrid methods}
\label{sec:Other hybrid methods}

Due to the complexity of the sketch recognition problem, more than one system proposed a hybrid method to solve those problems. They combine more than one approach to recognize symbols the user draws.
\cite{sketchunderstanding1,Simusketch51}  used four kind of classification image, graph matching, statistical and structure based classifiers. Each symbol will contain all graph, statistical, structure information and a (Wight, recognizer) tuples to determine which recognizer has the most influence in recognizing this symbol see \textit{Fig\ref{fig:structuralscemanticdeff}}.
\begin{figure}
	\centering
		\includegraphics[scale=0.5]{../../neededfiles/Figures/structuralscemanticdeff.eps}
	\caption[Structure definitions]{Example of the structure definition of sin symbol in \cite{sketchunderstanding1}.}
	\label{fig:structuralscemanticdeff}
\end{figure}

\cite {HMM53}  firstly normalize the stroke and compute all the statistical features from it. After that, a noise reduction algorithm is used on the strokes which will be translated to produce size and rotation invariance data. Those data is passed to a multistage hybrid recognizer. The first step of his multistage classifier is dynamic programming procedure to compute the optimal degree of similarity and differentiation of symbols. Secondly, a statistical feature ranking is employed on the data produced. Finally, the system will reject all unlikely classification and produce best fit for this stroke using a dynamic programming using domain and context knowledge as input.

\cite{statisticalparsing26} is a statically parser make use of the stored rules and probability of drawn stroked and uses Bayesian classifiers to do the classifications.

\subsection{AI based Classification}
\label{sec:AI based Classification}

Most system mentioned before use image or statistical feature based on the points the user draw. The next few sections will present methods that will use AI approached to classify the symbols the user draw.  Systems using AI approached imply a higher level of understanding of the strokes the user draws. Some construct a graph schema of the symbols from strokes the user draw. Other tried generating a visual grammar to identify the symbols that can be drawn.

\subsubsection{Graph Searching }
\label{sec:Graph Searching }

\textit{Figure \ref{fig:squarescematic}} shows a semantic graph that is constructed by the trainable recognizer in  \cite{mulitstroke5} for a rectangle. The trainable recognizer construct the semantic network at train time with then match the symbol user draw with the networks in the database. To match the networks quickly the system assume the user draws the symbol with the same order as in training set. A deep heuristic graph search will be employed if there was no match found.

@@

\cite{SRGraph57}  Uses two kind of graphs relations a spatial relationship graph (SRG) which is build on defining the spatial relationship between the source object and candidate object on the sketch. The relations are defined using Spatial Adjacency Relation (SAR) that falls into relation previously computed in system.   To make the graph fully descriptive the system computes Relative Position Relation of SAR relations and adds these relations into the SRG graph.

\subsubsection{Rule based }
\label{sec:Rule based }

The general algorithm of rule based system as used in \cite{interpretationmechanical50,SmartSketch56}  based on a knowledge of relation ship between nodes (strokes) a role is defined using those relation ships. The rule based system check the strokes the user draw against all strokes to detect if there a rule match the drawn symbol. When a rule matches the consequence of this rule, apply on the input symbol.

The rule system used in \cite{sketchinginterfaces2}  uses the (UI user interface) domain knowledge to do its basic recognition. It is divided into two step first is testing if the rules applies on the input symbol. Second step, is ranking the rules that apply on the symbol based on the highest confidence that was given to this input symbol.

Some system like in \cite{SmartSketch56}  used the rules to help in noise removal. Furthermore, rules based system are used to employ some reasoning to the recognized symbols with the help of context knowledge \cite {interpretationmechanical50}.

\subsubsection{HMM }
\label{sec:HMM }

 The main idea of HMM is representing knowledge and recognition in serious of states each state has its input conditions. Only a node that has the input conditions of the destination state that can transit it from one state to another. The final state is the recognition state of a given class. For example to recognize a rectangle the first state should be first line detected, second will be second line and the condition is the recognition of either a perpendicular line or parallel line to first one to another.

In sketch, recognition HMM was used to recognize symbols to different classes. Each class recognized was modeled using a single HMM model. The system in \cite {HMM53} encode the geometrical primitive that construct the symbol the user draws. After the encoding the codes are passed to HMM based classifier. Each object is represented by a HMM model and to manage multiple classes they combine matching score from different HMM models.

\subsubsection{Graph language and visual language describing }
\label{sec:Graph language and visual language describing }

Researchers tried to generate a visual language that can describe hand drawn sketches. In such systems, sketch recognition will be a question of parsing the sketch document into recognizer that parse strokes according to the constructed grammar.

Researchers proposed more than one visual grammar to use it in sketch recognition. \cite{statisticalparsing26} proposed a language parser that will parse sketched strokes using grammars rules that was constructed . The system first construct a parser that will parse drawn strokes into tokens then formulate those tokens using the grammar rules stored. Parsing is done in two steps; Firstly, the system should accept input of strokes then it use a statistical parser to generate the parse tree. The statically parser make use of the stored rules and probability of drawn stroked and uses Bayesian classifiers to produce the parse tree.

>>>>

\chapter{Literature Survey}
\label{sec:survey}
% this will contians only the finished work .... and my comments to view the old survey see the file
%draft_chaptersurvey 
%\section{Introduction }
%\label{sec:Introduction}
The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{sutherlandsketch}, which used a light pen to draw on the monitor to create circuit diagrams. Current technology and the increae in computuing power have increased the abilities of such interfaces and allowed the implementaion of fully sketching systems. Over the years, many other sketching based system have been developed, elthough throught this history most system foucused on building the interactive interfaces to assist users while entering information to the computer. To demonstrate our system inspiration we will first present current research in the area of sketch recognition. Fristly, preprosessing and segmentaiton methods will be discussed in details. Later, symbol recognition methods will be presented and analysied in Section \ref{sec:symbolrecogntion}.

% i need to write a history then small introudction to the survey...
%  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
\section{Preprocessing and Segmentation}
\label{sec:preprocessing}
%this section should describe how others did the first preprocessing of the problem. i have to summarize what i wrote before. 
 % first paragrah about how stroke is a messing thing and naturally it is ambigues. 
  In a system that supports free hand sketching, sketching should be as natural as possible. This means that the user can draw symbols using single or multiple strokes without constrains. The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper. But sketchs are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users. This led researchers to perform sort of preprocessing before starting the sketch recognition task itself. The next sections explain different preprocessing techniques.

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/Overshootandundershoot.eps}
	\caption[Neat and Sloppy symbols] {A sloppy and neat versions of symbol.}
	\label{fig:Overshootandundershoot}
\end{figure}

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/overstroked.eps}
	\caption{Example of an over traced stroke}
	\label{fig:overstroked}
\end{figure}

 %The preprocessing can be noise removing, stroke beautification or segmentation.  %Because of this 
 Sezgin and Davis \cite{overtraced24} tried to solve the over traced strokes (see Figure \ref {fig:overstroked})using a thinning algorithm. In \cite{meanshift10,domainindependent17}, Bo Yu uses mean shift procedure as a noise remover and preprocessing step to the recognition step. The mean shift procedure is simply shifting the point to the average of its neighborhoods so smoothing out the noise without changing the global structure of data. Gaussian filters was used to smooth curves \cite {Phoenix88}. These procedures may seem to remove noises in the sketch but it often treat vertices as noise which obscure vertex locations. 
 %% now the third paragraph.....
Beside points location information (x, y positions) researcher made uses of the extra information from using digital ink. Some as \cite{mulitstroke5,polygonfeedback31} used user speed information to detect critical points where it is stated that user slow down at those critical points. Other used curvature and directions data. %Others \cite {MilitaryCOA37,computationalmodel16,polygonfeedback31}, debated on the importance of the extra information and weather the temporal and speed information is significant.

 Agar et al. \cite{polygonfeedback31} used the timestamp for each point and compute time difference between every two successive point to help him finding corners as it provides more distinct maxima than the speed information. He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals while the pointing device is moving. There are no points while the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary. Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data.

Sezgin et al.\cite{earlyprocess} used both speed and direction information. Their system gives better results than using only speed or direction information. This is because the speed information tend to be noisy when drawing large symbols, on the other side when using only direction information the some corner are missed. Our system uses speed, direction and time difference information in its preliminary calculation step.  %both speed and direction information help detecting vertices when the

 A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al. \cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system. Also, they did not take advantage of the curvature or local geometric properties of the digital curve. Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts  Yin \cite{PolygonApproximationPSO} method and tries to improve it by adding curvature and other local information while segmenting strokes to achieve better segmentations. %The next section presents the general particle swarm algorithm which is used in both our segmentation algorithms. % are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}

  % third should speech on how systems define critical points (this paragrah should speak on the data collected (speed crital points, ...

 After critical point detection most sketch system proceeds to label segments into simple geometric primitives (i.e line, arc, etc...).  A hybrid algorithm was introduced in \cite{earlyprocess} where different sets of segments are generated based on both curvature and speed dominant points, followed by choosing a segmentation with the least error from a generated hybrid set. However, this system is limited to recognizing only specific simple geometric shapes with a set of low level recognizers. Each low level recognizer is designed to recognize only one geometric shape using spatial and geometric information extracted from input stroke.  
Yu \cite{meanshift10} introduced a \textit{feature area} for each primitive and then computed the segmentation error for different types of primitives based on the computed \textit{feature area}. Figure \ref{fig:featurearearc} shows the feature area of a line and circle in Yu system, the shaded area is the difference between the computed feature area and the assumption area. His system achieved good accuracy in simple shapes (square, ellipse,...etc) but did not perform well in complex shapes.

 Paulson and Hammond \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\% accuracy. However, their system similar to all low level recognizers in the way that it identifies a small set of simple shapes.% A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.
\cite{Fluid25} Trying to combine fast morphing and basic recognition for the sketch drawn as the user draw it. It tries to move the points of strokes drawn by the user draw into one of the known shapes (ex: box, lines and circle). It uses a relaxation and least square method for circles and uses the string of forces for finding best-fit box. However, their system can only identify circle and rectangles drawn using a single stroke.  
\begin{figure}
	\centering
	\subfigure[Feature area of line an arc]{
		\includegraphics{images/featurearearc.eps}}
		\subfigure[Feature area of line]{
				\includegraphics{images/featureareline.eps}}
	\caption{Feature area}
	\label{fig:featurearearc}
\end{figure}

\section{Symbol Recogntion}
\label{sec:symbolrecogntion}

In this section, we review three different types of symbol recognition methods. Symbol recognition can be based on: strokes, global shape properties, and appearance \cite{Oltmans07}. The role of which each stroke plays in the structure of the sketched symbol is used to determine shapes in stroke based methods. In global shape properties approach, systems investigate general shape properties and its underlying strokes. Using shape properties relaxes the assumption and role of each stroke in the symbol but it does not truly represent the appearance of the symbol. Unlike appearance based methods where individual strokes are disregarded and they focus on the overall appearance of shapes. We choose to use the appearance based method but added some geometrical and structure properties used in most stroke based approaches. Our method tries to avoid many of the problems faced by approaches based on individual strokes or global shape properties while trying to gain the additional information that the stroke based system use. In the next sections, we look into three different symbol recognition approaches, describe their advantages and challenges.

\subsection{Stroke Based Recognition}
\label{sec:StrokebasedRecognition}
Stroke based approach is based on the assumption that each stroke has a specific role in representing the symbol. Stroke based methods examine each stroke as the user draw it to determine its role in the symbol. These methods are suited to interactive interfaces since users expected that system displays interpretation of the symbol after the stroke is drawn or after drawing a group of strokes. The majority of the research in sketch recognition has focused on stroke-based methods because of the corresponding emphasis on interactive interfaces. Most stroke based systems either represent symbols as gestures or   hierarchical shape.

\subsubsection{Guesture Recognition}
\label{sec:GuestureRecognition}

Single stroke gestures recognition was the focus of the early work in symbol recognition \cite{gestureexample12,aideddesgin22,sketchinginterfaces2}.  Gestures are pen strokes drawn by the user that can immediately recognized by the system. The result was either creating an action (e.g. deleting a shape) or drawing a shape on the screen. These systems are based on how the user draw the shape rather than how it looks.

One of the first systems in gestures recognition is Rubine's algorithm \cite{gestureexample12} . He used Gaussian statistical classifier to classify gestures.  Simple features as length of stroke, smoothness of stroke, the angle that the first part of the gesture is drawn at and properties of the bounding box of the stroke are used to train the classifier to recognize  symbols. In our system, we combine Rubine features with others to include the information on how the user draws sketch while recognizing the symbols. In \cite{Long00}  Long shows how similarities between gestures can be analyzed.  He used this information to identify gestures that are likely to confuse the recognition system. This helped interface designers to choose gestures that are easily recognized.

The main problem with gesture based recognition system that they impose several constrains on how the user can draw. Gestures must be drawn using a specific order and direction to be recognized correctly. Moreover, in most cases gestures do not necessarily look like the symbols they represent (e.g. to represent a rectangle the user draws only the left edge of it). Because of these reasons, gestures recognition systems are not suitable for free hand drawn sketches.

\subsubsection{Hierarchical Shape Descriptions}

A Hierarchical shape descriptions was widely used in sketch systems \cite{sketchunderstanding1,HierarchicalParsing7,Ladder30,AlvaradoFreedom42,napkinGross}. Symbols are represented using hierarchical levels. The lower levels represent simple geometric primitives such as lines, arc and ellipses. Intermediate levels shapes are composed of lower level parts and constrain or connections between them. For example, a triangle is described as three lines connected at their endpoints. Higher level shapes are constructed from intermediate and lower level shapes, for example, a diode is described as a line on the top of a triangle head (acute angle).

 Systems that employ such representation must first preprocess strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods). Recognition can then be treated as sub graph matching problem.

%Although not based on a formal structural model of shapes, another probabilistic approach by Szummer and Qi in [37] uses conditional random fields (CRFs) to propagate information about the labeling of one stroke fragment to its neighbors. This allows their system to assign labels to stroke fragments that cannot easily be identified in isolation and to regroup over-segmented fragments. This helps mitigate the difficulty of determining the exact granularity at which to perform fragmentation by using the context of surrounding fragments and a global interpretation of the sketch. They have applied their algorithm to the binary labeling problem of distinguishing boxes from connectors in organizational charts with good results, even in several highly ambiguous cases.
\paragraph{Sub-graph matching:}
\label{sec:GraphSearching}

Representing the symbol that the user draws as a graph transforms the classification problem into a graph-matching problem. In general, each user geometrical primitive is represented as a node of the graph and the relations between those nodes are represented as arc.  Figure \ref{fig:squarescematic} shows a graph that is constructed by \cite{mulitstroke5} for a rectangle. To classify the symbols the graph constructed must be matched with the graphs in the database.

@@

In the worst case, sub-graph matching is an exponential problem and thus is expensive to compute. Researcher tried to reduce complexity by restricting the search to fragments that are spatially or temporally grouped, or by using other assumption. Some of the common assumptions are that each stroke is only part of one symbol and that user must draw one shape before drawing the next\cite{physicalmeaning6}. These assumptions add constrains on how user can draw sketches and thus violates freely drawn sketches.  Moreover, the matching process in highly affected by segmentation results. Over or under fragmentation of stroke into geometric primitive complicates the matching process.  If a stroke is broken into too many fragments then there will be extra components that will not map to any part of the symbol. If the stroke is divided enough then there will be component of the shape that are not filled while matching. The different ways of segmenting the sketches worsen the already high cost graph matching.

 To solve this problem,  \cite{mulitstroke5} provided two methods to match graphs, a simpler one using restrictive assumptions and if it fails the system uses a longer more computationally expensive search.  . Figure \ref{fig:squarescematic} shows a semantic graph that is constructed by the trainable recognizer in   \cite{mulitstroke5} for a rectangle. The trainable recognizer construct the semantic network at train time with then match the symbol user draw with the networks in the database. To match the networks quickly the system assumes the user draws the symbol with the same order as in training set. A deep heuristic graph search will employed only if no match was found. % there was no match found.

 Several other systems tried to model the matching process as a probabilistic problem \cite{HMM53,SketchRead2007}. These systems allow matching of partially satisfied shapes to propagate down to re interpret low level fragmentation hypothesis. In SKETCREAD \cite{SketchRead2007}, a dynamically generated Bayesian network fragments represents the shape hypotheses. Higher level structure of shape fragments can cause re interpretation of the geometric primitives. For example, if one head of an arrow is initially labeled as an arc and the system fragments the shape as an arrow then the belief that this stroke is an arc is decrease and the belief that it is a line increase. Finally, the arrow structure is completed and it is fully recognized.

 \paragraph{Visual Describtive Languages:}
 \label{sec:VisualDescribtiveLanguages}

 Avoiding graph matching problems, research tried to use languages that discribe shapes instead of graphs. Visual describtive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives. Figures \ref{fig:arrow1} and \ref{fig:arrow2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}. Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.
